---
title: "Copernicus_Med_MHW"
author: "Dan Bez Golanski"
date: "`r Sys.Date()`"
output: html_document
---
## Load packages
```{r warning=FALSE}
library(ncdf4) # for loading copernicus data
library(lubridate)
library(tidyverse)
library(heatwaveR)
library(openxlsx)
library(sjPlot) # for liner detrending model
library(ggpubr) # for linear regression equation and R squared in the IUI and Copernicus 
library(zoo) # for moving median
library(forecast)
library(tidync) # For easily dealing with NetCDF data
library(rerddap) # For easily downloading subsets of data
library(doParallel) # For parallel processing

```
## Functions
```{r}
sst_copernicus_to_df <- function(coperniucs_nc,names_vector,depth,Celsuis) {
dim_lon <- ncvar_get(coperniucs_nc,names_vector[1])
dim_lat <- ncvar_get(coperniucs_nc,names_vector[2])
dim_time <- ncvar_get(coperniucs_nc,names_vector[3])
t_units <- ncatt_get(coperniucs_nc,"time","units") 
t_ustr <- strsplit(t_units$value, " ")
t_dstr <- strsplit(unlist(t_ustr)[3], "-")
sst <- ncvar_get(coperniucs_nc,names_vector[4],collapse_degen = F)
# analysis_error <- ncvar_get(coperniucs_nc,"analysis_error",collapse_degen = F)
sst_wide <- as.matrix(expand.grid(dim_lon,dim_lat))
sst_wide <- as.data.frame(sst_wide)
names(sst_wide) <- c("lon","lat")
sst_wide <- sst_wide %>% arrange(lon)
for (year_i in 1:length(dim_time)) {
  if (depth) temp_date <-sst[,,,year_i] else temp_date <-sst[,,year_i] #If more than one depth this row need to be changed 
  temp_date <- t(temp_date)
  vec_temp_date <- c(temp_date)
  sst_wide[,year_i+2] <- vec_temp_date
}
names(sst_wide) <- c("lon","lat",dim_time)
sst_long <- cbind(sst_wide[1:2], stack(sst_wide[3:dim(sst_wide)[2]]))
names(sst_long) <- c("lon","lat","sst","date")
sst_long <- na.omit(sst_long)
sst_long$date <- as.numeric(as.character(sst_long$date))
sst_long$date <- ymd(t_dstr)+ dseconds(sst_long$date)
if (!Celsuis) sst_long$sst <- sst_long$sst - 273.15
return(sst_long)
}

Detrend_MHW_days_to_MHWs <- function(deterend_df, date_col_num, MHW_day_col_num, residuals_col_num, origin_col_num, threshold_col_num) {
  
# Initialize an empty data frame to store MHWs data with various attributes
MHWs_df <- data.frame(date_start=NA,
                        date_end=NA,
                        date_peak=NA,
                        duration=NA,
                        intensity_mean=NA,
                        intensity_max=NA,
                        intensity_var=NA,
                        intensity_cum=NA,
                        intensity_mean_abs=NA,
                        intensity_max_abs=NA,
                        intensity_var_abs=NA,
                        intensity_mean_relThresh=NA,
                        intensity_max_relThresh=NA,
                        intensity_var_relThresh=NA,
                        rate_onset=NA,
                        rate_decline=NA
                        )

  # Create lists to store information about MHWs
  resid_mhw <- list()       # Stores the residual values during the MHW event
  absolute_mhw <- list()    # Stores the original values during the MHW event
  thresh_mhw <- list()      # Stores the threshold values during the MHW event
  
  # Variable to store the start date of an MHW event
  date_start <- NA
  for (day_i in 1:length(deterend_df[, date_col_num])) {
    
    # If it's the start of a new MHW event
    if(deterend_df[day_i, MHW_day_col_num] == TRUE && is.na(date_start)) {
      date_start <- deterend_df[day_i, date_col_num]  # Set the start date of the event
      resid_mhw <- append(resid_mhw, deterend_df[day_i, residuals_col_num])  # Append the residual value
      absolute_mhw <- append(absolute_mhw, deterend_df[day_i, origin_col_num])  # Append the original value
      thresh_mhw <- append(thresh_mhw, deterend_df[day_i, residuals_col_num] - deterend_df[day_i, threshold_col_num])  # Append the threshold difference
      mhw_gap <- 0  # Initialize gap counter
      
    } 
    
    # If the event continues (another day marked as MHW)
    else if(deterend_df[day_i, MHW_day_col_num] == TRUE && !is.na(date_start)) {
      resid_mhw <- append(resid_mhw, deterend_df[day_i, residuals_col_num])
      absolute_mhw <- append(absolute_mhw, deterend_df[day_i, origin_col_num])
      thresh_mhw <- append(thresh_mhw, deterend_df[day_i, residuals_col_num] - deterend_df[day_i, threshold_col_num])
      mhw_gap <- 0
      
    } 
    
    # If the event has ended and the gap exceeds 2 days
    else if (deterend_df[day_i, MHW_day_col_num] == FALSE && !is.na(date_start)) {
      mhw_gap <- mhw_gap + 1  # Increase the gap counter
      
      # If the gap exceeds 2 days, the MHW event is considered finished
      if(mhw_gap > 2) {
        
        # If the MHW event lasted more than 5 days, calculate various metrics and store them
        if(length(resid_mhw) > 5) {
          temp_df <- data.frame(date_start = date_start,
                                date_end = deterend_df[day_i - 1, date_col_num],  # End date is the last day of the event
                                date_peak = date_start + which.max(unlist(resid_mhw) - 1),  # Peak date based on maximum residual
                                duration = deterend_df[day_i, date_col_num] - date_start,  # Duration of the event
                                intensity_mean = mean(unlist(resid_mhw)),  # Mean intensity of the MHW event
                                intensity_max = max(unlist(resid_mhw)),  # Maximum intensity of the MHW event
                                intensity_var = sd(unlist(resid_mhw)),  # Standard deviation of intensity
                                intensity_cum = sum(unlist(resid_mhw)),  # Cumulative intensity of the event
                                intensity_mean_abs = mean(unlist(absolute_mhw)),  # Mean absolute intensity
                                intensity_max_abs = max(unlist(absolute_mhw)),  # Maximum absolute intensity
                                intensity_var_abs = sd(unlist(absolute_mhw)),  # Standard deviation of absolute intensity
                                intensity_mean_relThresh = mean(unlist(thresh_mhw)),  # Mean intensity relative to the threshold
                                intensity_max_relThresh = max(unlist(thresh_mhw)),  # Maximum intensity relative to the threshold
                                intensity_var_relThresh = sd(unlist(thresh_mhw)),  # Standard deviation of intensity relative to threshold
                                rate_onset = max(unlist(resid_mhw)) / which.max(unlist(resid_mhw)),  # Rate of onset (intensity over time)
                                rate_decline = max(unlist(resid_mhw)) / if(length(resid_mhw) - which.max(unlist(resid_mhw)) == 0) 1 else (length(resid_mhw) - which.max(unlist(resid_mhw)) + 1)  # Rate of decline
          )
          
          # Append the calculated MHW event data to the MHWs_df data frame
          MHWs_df <- rbind(MHWs_df, temp_df)
        }
        
        # Reset the variables for the next MHW event
        mhw_gap <- 0
        resid_mhw <- list()
        absolute_mhw <- list()
        thresh_mhw <- list()
        date_start <- NA
      }
    }
  }

  # If any rows were added to MHWs_df, remove the initial NA row
  if(dim(MHWs_df)[1] > 0) {
    MHWs_df <- MHWs_df[-1,]
  }

  # Convert date columns to Date type
  MHWs_df$date_start <- as.Date(MHWs_df$date_start)
  MHWs_df$date_end <- as.Date(MHWs_df$date_end)
  MHWs_df$date_peak <- as.Date(MHWs_df$date_peak)

  # Return the MHW events data frame
  return(MHWs_df)
}
Draw_sst_time <- function(sst_df,lat_f=NA, lon_f=NA,one_cor=T) {
  if(one_cor)
  {
    sst_df <- sst_df %>% filter(lat == sst_df$lat[1] & lon == sst_df$lon[1])
  } else if (is.na(lat_f) | is.na(lon_f)) {
    stop("lat or lon is missing")
  } else
  {
  sst_df <- sst_df %>% filter(lat == lat_f & lon == lon_f)
  }
  ggplot()+
  geom_point(aes(x=sst_df$date,sst_df$sst)) +
  ylab("SST")+
  xlab("Date")
  
}
OISST_sub_dl <- function(time_df){
  OISST_dat <- griddap(datasetx = "ncdcOisst21Agg_LonPM180", 
                       url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                       time = c(time_df$start, time_df$end), 
                       zlev = c(0, 0),
                       latitude = c(29.472, 29.578),
                       longitude = c(34.894, 35.017),
                       fields = "sst")$data %>% 
    mutate(time = as.Date(stringr::str_remove(time, "T00:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst,lon = longitude, lat = latitude) %>% 
    select(lon, lat, t, temp) %>% 
    na.omit()
}

```

## Load data
```{r}

# Med Sea
med_sst_long<- readRDS("Med/Israel Climtology/Coper_SST_Med.RDS")
territorial_water_coord_med_cop <-read.csv("Med/Community Ecology/Isr_teritorial_water_locations.csv") 
territorial_water_coord_med_cop <- territorial_water_coord_med_cop %>% select(Station,Longitude.,Latitude.d)

# Red Sea
#daily data from web with avg between max and min from continuous data
sst_Eilat_IUI_daily <- read.xlsx("Red Sea/Eilat climetology/IUI/Eilat Daily Metro_fix_outliers.xlsx")
sst_Eilat_IUI_daily <- sst_Eilat_IUI_daily[-1:-4,]
colnames(sst_Eilat_IUI_daily) <- c("num_date","Temp_max","Temp_min","Temp_avg")
sst_Eilat_IUI_daily$Date <- as.Date(as.numeric(sst_Eilat_IUI_daily$num_date),origin = "1899-12-30")
sst_Eilat_IUI_daily$Temp_avg <- as.numeric(sst_Eilat_IUI_daily$Temp_avg)

sst_IUI_10_min <- read.csv("Red Sea/Eilat climetology/IUI/SST_Eilat_IUI_10min.csv")
sst_IUI_1day <- readRDS("Red Sea/Eilat climetology/IUI/SST_Eilat_IUI_1day.RDS")# this dataset is derived from the 10 min dataset and is calculated in "IUI 10min format handling and convert to 1day" section

OISST_data_red <- readRDS("Red Sea/Eilat climetology/OISST_SST_Red.RDS")

IUI_amatzia_daily_temp <-read.xlsx("Red Sea/Eilat climetology/Amazia - IUI/Obs Pier SST measured and interpolated 1988-2023.xlsx",sheet = "Interpolation") 

```

## Load ncdf
```{r}
med_coperniucs_nc <- nc_open("Med/Israel Climtology/cmems_SST_MED_SST_L4_REP_OBSERVATIONS_010_021_1732629959487.nc")
#OSTIA
red_copernicus_nc <- nc_open("Red Sea/Eilat climetology/METOFFICE-GLO-SST-L4-REP-OBS-SST_1705328292068.nc")
#ESA & CCI
red_copernicus_nc_2017 <- nc_open("Red Sea/Eilat climetology/ESACCI-GLO-SST-L4-REP-OBS-SST_1711473070752.nc")
red_copernicus_nc_2022 <- nc_open("Red Sea/Eilat climetology/C3S-GLO-SST-L4-REP-OBS-SST_1711473119162.nc")
#Physics
red_copernicus_nc_physics <- nc_open("Red Sea/Eilat climetology/cmems_mod_glo_phy_my_0.083deg_P1D-m_1711477180532.nc")
noaa_sst <- nc_open("../../NOAA OISTT v2/subset_try.nc")
```

## Get OISST
```{r}
rerddap::info(datasetid = "ncdcOisst21Agg_LonPM180", url = "https://coastwatch.pfeg.noaa.gov/erddap/")

dl_years <- data.frame(date_index = 1:5,
                       start = c("1982-01-01", "1990-01-01", 
                                         "1998-01-01", "2006-01-01", "2014-01-01"),
                       end = c("1989-12-31", "1997-12-31", 
                                       "2005-12-31", "2013-12-31", "2022-12-31"))
# Download all of the data with one nested request
# The time this takes will vary greatly based on connection speed
system.time(
  OISST_data <- dl_years %>% 
    group_by(date_index) %>% 
    group_modify(~OISST_sub_dl(.x)) %>% 
    ungroup() %>% 
    select(lon, lat, t, temp)
) # 38 seconds, ~8 seconds per batch
saveRDS(OISST_data,"Red Sea/Eilat climetology/OISST_SST_Red.RDS")
```
## Create dataframe from copernicus nc
```{r}
med_sst_long <- sst_copernicus_to_df(med_coperniucs_nc,c("longitude","latitude","time","analysed_sst"),F,F)
#Physics

saveRDS(med_sst_long,"Coper_SST_Med.RDS")
```


## IUI 10min format handling and convert to 1day
```{r}
sst_IUI_10_min$Average.temp <- as.double(sst_IUI_10_min$Average.temp)
sst_IUI_10_min <- sst_IUI_10_min %>% select(DateTime,Average.temp)
sst_IUI_10_min <- na.omit(sst_IUI_10_min)
sst_IUI_10_min$excel <- parse_date_time(sst_IUI_10_min$DateTime,c("%d/%m/%Y %H:%M:%S", "%d/%m/%Y %H:%M"))
sst_IUI_10_min$day <- cut(sst_IUI_10_min$excel,breaks = "1 day")
sst_IUI_1day <- sst_IUI_10_min %>% 
  select(excel,day,Average.temp) %>% 
  group_by(day) %>% 
  summarise_at(vars(Average.temp) ,funs(mean,n()))
missing_days <- 0
sst_IUI_1day$day <- as.Date(sst_IUI_1day$day)

for (day_i in 2:nrow(sst_IUI_1day)) {
      temp_diff <-as.numeric(difftime(sst_IUI_1day$day[day_i],sst_IUI_1day$day[day_i-1],units = 
                                        "days"))
      if (temp_diff>1) {
        missing_days <- missing_days + temp_diff
      }
}
print(paste("There are",missing_days,"missing days."))
ggplot(sst_IUI_1day)+
  geom_point(aes(x=day,y=mean))+
  scale_x_date(breaks = "1 year")+
  xlab("Date")+
  ylab("Avg daily temp")
saveRDS(sst_IUI_1day,"Red Sea/Eilat climetology/IUI/SST_Eilat_IUI_1day.RDS")
## There are 43 missing days in this database
```

## Calculate climatelogy for whole med and MHW's events
```{r} 

coordinates_list <- med_sst_long %>% filter(date==med_sst_long$date[1])
coordinates_list <- coordinates_list %>% select(lat,lon)
MHW_med <- data.frame(lat=double(),
                      lon=double(),
                      date_start=Date(),
                      date_end=Date(),
                      duration=double(),
                      intensity_mean=double(),
                      intensity_max=double(),
                      intensity_cumulative=double(),
                      rate_onset=double(),
                      rate_decline=double()
)
for (coord_num in 1:dim(coordinates_list)[1]) {
  temp_point <- med_sst_long %>% filter(lat==coordinates_list$lat[coord_num] &coordinates_list$lon[coord_num])
  #deleting lat and lon cause all in same position
  temp_point <- temp_point[-1:-2]
  temp_point <- temp_point %>% arrange(date)

  names(temp_point) <- c("temp","t")
  clim_period <- c(temp_point$t[1],"2011-08-25")
  MHW_climatelogy_tresh <- ts2clm(temp_point,climatologyPeriod = clim_period,pctile = 90)
  # MCS_climatelogy_tresh <- ts2clm(temp_point,climatologyPeriod = clim_period,pctile = 10)
  MHW_events <- detect_event(MHW_climatelogy_tresh)
  MHW_events <- MHW_events$event %>%    select(duration,date_start,date_end,intensity_mean,intensity_max,intensity_cumulative,rate_onset,rate_decline)
  # MCS_events <- detect_event(MCS_climatelogy_tresh,coldSpells = T)
  MHW_events$lat <- coordinates_list$lat[coord_num]
  MHW_events$lon <- coordinates_list$lon[coord_num]  
  MHW_med <- rbind(MHW_med,MHW_events)
  print(coord_num)
}
MHW_med$year <- format(MHW_med$date_start,"%Y")
saveRDS(MHW_med,"MHW_Events_Isr_Med_allpoints.RDS")
write.csv(MHW_med,"MHW_Events_Isr_Med_allpoints.csv")

```


## Calculate climatelogy for specific point 
```{r}
# Set playground coordinates (latitude and longitude for a specific location).
playground_coord <- c(32.38210, 34.82362)

# Filter the Mediterranean Sea SST dataset for the specified coordinates.
# Round latitudes and longitudes to 5 decimal places to  match them with playground_coord.
med_sst_coord <- med_sst_long[
  round(med_sst_long$lat, 5) == playground_coord[1] & 
  round(med_sst_long$lon, 5) == playground_coord[2],
]

# Remove latitude and longitude columns from the filtered data
# since all values correspond to the same location.
sst_coord <- med_sst_coord %>% select(-lat, -lon)

# Arrange the SST data by date for proper temporal analysis.
sst_coord <- sst_coord %>% arrange(date)

# Rename the columns for consistency and clarity:
# "temp" represents sea surface temperature (SST), and "t" represents the date.
names(sst_coord) <- c("temp", "t")

# Define the climatology period for SST analysis.
# The climatology period starts from the first date in the data and preferably ends 30 years later.
clim_period <- c(sst_coord$t[1], "2011-08-25")

# Compute the climatology and the 90th percentile threshold for marine heatwave (MHW) detection.
MHW_climatelogy_tresh <- ts2clm(sst_coord, climatologyPeriod = clim_period, pctile = 90)

# Optional: Uncomment this section to calculate thresholds for marine cold spells (MCS).
# MCS_climatelogy_tresh <- ts2clm(sst_coord, climatologyPeriod = clim_period, pctile = 10)

# Detect marine heatwave (MHW) events based on the climatology and threshold.
MHW_events <- detect_event(MHW_climatelogy_tresh)

# Optional: Uncomment this section to detect marine cold spell (MCS) events.
# MCS_events <- detect_event(MCS_climatelogy_tresh, coldSpells = TRUE)

# Categorize detected MHW events.
MHW_events_cat <- category(MHW_events)

# Rename the peak date column for clarity and consistency.
MHW_events_cat <- MHW_events_cat %>% rename("date_peak" = "peak_date")

# Extract the climatology data for visualization or further analysis.
climatelogy <- MHW_events$climatology

# Select relevant columns from the detected MHW events.
MHW_events <- MHW_events$event %>% select(-index_start, -index_peak, -index_end)

# Merge the categorized event information (e.g., intensity category) with the main MHW events data.
MHW_events <- merge(MHW_events, MHW_events_cat[, c("event_no", "category")], by = "event_no")

# Add the latitude and longitude of the playground location to each MHW event.
MHW_events$lat <- playground_coord[1]
MHW_events$lon <- playground_coord[2]

# Extract the year and month from the event start date for temporal analysis.
MHW_events$year <- format(MHW_events$date_start, "%Y")
MHW_events$month <- format(MHW_events$date_start, "%m")

# Create a histogram of the months with the most MHW events.
# The histogram focuses on short-duration events (less than 31 days) after certain year.
hist(as.numeric(MHW_events$month[MHW_events$duration < 31 & MHW_events$year > 2014]))

```

## MHW Isr_med_statistics_allpoints
```{r}
#Statistics for Israeli coast 
MHW_med_sum_year_date_start <- MHW_med %>% group_by(year,date_start,duration) %>% summarise(Coord_num=n(),avg_mean_intensity=mean(intensity_mean),avg_max_intensity=mean(intensity_max),avg_intensity_cummelative=mean(intensity_cumulative),avg_onset_rate=mean(rate_onset),avg_decline_rate=mean(rate_decline))
MHW_med_sum_year_date_start$area_ratio <- MHW_med_sum_year_date_start$Coord_num/1443
MHW_med_spatial_filter <- MHW_med_sum_year_date_start[MHW_med_sum_year_date_start$area_ratio>0.2,]
write.csv(MHW_med_spatial_filter,"Isr_MHW_Med_0.2_area_filter.csv")

```


## Compute marine heat waves for the territorial israeli coast mean sst
```{r}
coordinates_list <- med_sst_long %>% filter(date==med_sst_long$date[1])
med_sst_long$lat <- round(med_sst_long$lat,4)
med_sst_long$lon <- round(med_sst_long$lon,4)
territorial_water_coord_med_cop$Longitude. <- round(territorial_water_coord_med_cop$Longitude.,4)
territorial_water_coord_med_cop$Latitude.d <- round(territorial_water_coord_med_cop$Latitude.d,4)
territorial_water_coord_med_cop$coord <- paste(territorial_water_coord_med_cop$Longitude.,territorial_water_coord_med_cop$Latitude.d)
med_sst_long$coord <- paste(med_sst_long$lon,med_sst_long$lat)
med_ter_ISR_sst <- med_sst_long %>% filter(coord %in% territorial_water_coord_med_cop$coord)
#The result of the filtering missing 14 points from two different longitudes : 34.52335 & 35.07385
mean_Isr_coast <- med_ter_ISR_sst %>%
  group_by(date) %>%
  summarise(mean_sst=mean(sst))

clim_period <- c(mean_Isr_coast$date[1],"2011-08-25")
names(mean_Isr_coast) <- c("t","temp")
MHW_climatelogy_tresh <- ts2clm(mean_Isr_coast,climatologyPeriod = clim_period,pctile = 90)
# MCS_climatelogy_tresh <- ts2clm(temp_point,climatologyPeriod = clim_period,pctile = 10)
MHW_events <- detect_event(MHW_climatelogy_tresh)
event_line(MHW_events)
MHW_events <- MHW_events$event %>%    select(duration,date_start,date_end,date_peak,intensity_mean,intensity_max,intensity_cumulative,rate_onset,rate_decline)
write.csv(MHW_events,"Med/Community Ecology/MHW_Isr_mean_territorial_waters.csv")
```

## Calculate climatelogy for red sea(OISST) and MHW's events
```{r} 
OISST_data_red <- OISST_data_red %>%  select(-lat,-lon)
OISST_data_red <- OISST_data_red %>% arrange(t)
clim_period <- c(OISST_data_red$t[1],as.Date("2011-12-31"))#edit date to 30 years from beginning if data is available
MHW_climatelogy_tresh <- ts2clm(OISST_data_red,climatologyPeriod = clim_period,pctile = 90)
MHW_events <- detect_event(MHW_climatelogy_tresh)
# MCS_climatelogy_tresh <- ts2clm(red_sst_long,climatologyPeriod = clim_period,pctile = 10)
# MCS_events <- detect_event(MCS_climatelogy_tresh)
# MCS_events <- MCS_events$event
MHW_events_cat <- category(MHW_events)
MHW_events_cat <- MHW_events_cat %>% rename("date_peak" = "peak_date")
climatelogy <- MHW_events$climatology
MHW_events <- MHW_events$event %>% select(-index_start, -index_peak, -index_end)
MHW_events <- merge(MHW_events,MHW_events_cat[,c("event_no","category")],by="event_no")
hist(MHW_events$intensity_mean)
# MCS_events <- detect_event(MCS_climatelogy_tresh,coldSpells = T)
excel <- createWorkbook()
addWorksheet(excel,sheetName = "MHW_Events")
addWorksheet(excel,sheetName = "Climatelogy")
writeData(excel,"MHW_Events",MHW_events)
writeData(excel,"Climatelogy",climatelogy)
saveWorkbook(excel,"Red Sea/Eilat 2017 MHW/MHWs/MHW_OISST_Events_RedSea_Fix")#add description to name according to database used

```
## Calculate climatelogy for red sea(Amatzia) and MHW's events
```{r} 
IUI_amatzia_daily_temp$date <- as.Date(IUI_amatzia_daily_temp$Doy,origin = as.Date(paste(IUI_amatzia_daily_temp$Year,"-01-01",sep = ""))-1)
IUI_amatzia_daily_temp <- IUI_amatzia_daily_temp %>%
  select(date,SST) %>% 
  rename("t"="date","temp"="SST")
clim_period <- c(IUI_amatzia_daily_temp$t[1],as.Date("2015-12-31"))#edit date to 30 years from beginning if data is available
MHW_climatelogy_tresh <- ts2clm(IUI_amatzia_daily_temp,climatologyPeriod = clim_period,pctile = 90)
MHW_events <- detect_event(MHW_climatelogy_tresh)
# MCS_climatelogy_tresh <- ts2clm(red_sst_long,climatologyPeriod = clim_period,pctile = 10)
# MCS_events <- detect_event(MCS_climatelogy_tresh)
# MCS_events <- MCS_events$event
MHW_events_cat <- category(MHW_events)
MHW_events_cat <- MHW_events_cat %>% rename("date_peak" = "peak_date")
climatelogy <- MHW_events$climatology
MHW_events <- MHW_events$event %>% select(-index_start, -index_peak, -index_end)
MHW_events <- merge(MHW_events,MHW_events_cat[,c("event_no","category")],by="event_no")
hist(MHW_events$intensity_mean,breaks = seq(from=0.4, to=3, by=0.2))
# MCS_events <- detect_event(MCS_climatelogy_tresh,coldSpells = T)
excel <- createWorkbook()
addWorksheet(excel,sheetName = "MHW_Events")
addWorksheet(excel,sheetName = "Climatelogy")
writeData(excel,"MHW_Events",MHW_events)
writeData(excel,"Climatelogy",climatelogy)
saveWorkbook(excel,"Red Sea/Eilat 2017 MHW/MHWs/MHW_Amatzia_Events_RedSea_Fix.xlsx")#add description to name according to database used

```
## Calculate climatelogy for Red(IUI) and MHW's events
```{r}
names(sst_IUI_1day) <-c("t","temp","n") 
sst_IUI_1day <- sst_IUI_1day[,c(1,2)]
clim_period <- c(sst_IUI_1day$t[1],as.Date("2015-12-31"))
MHW_climatelogy_tresh <- ts2clm(sst_IUI_1day,climatologyPeriod = clim_period,pctile = 90)
# MCS_climatelogy_tresh <- ts2clm(temp_point,climatologyPeriod = clim_period,pctile = 10)
MHW_events <- detect_event(MHW_climatelogy_tresh)
climatelogy <- MHW_events$climatology
MHW_events <- MHW_events$event %>%    select(date_start,date_end,duration,intensity_mean,intensity_max,intensity_cumulative,rate_onset,rate_decline)
# MCS_events <- detect_event(MCS_climatelogy_tresh,coldSpells = T)
excel <- createWorkbook()
addWorksheet(excel,sheetName = "MHW_Events")
addWorksheet(excel,sheetName = "Climatelogy")
writeData(excel,"MHW_Events",MHW_events)
writeData(excel,"Climatelogy",climatelogy)
saveWorkbook(excel,"Red Sea/Eilat 2017 MHW/MHWs/MHW_IUI_1Daymean_Events_RedSea_Fix.xlsx")
```



## Compare IUI and Coper/OISST data
```{r}
compare_sst <- OISST_data_red
compare_sst$year <- year(compare_sst$t)
compare_sst <- compare_sst %>% filter(year>=2006)
compare_sst$Date <- as.Date(compare_sst$t)
sst_IUI_1day <- sst_IUI_1day %>% rename("Date"="day")
#merge dataset  mean_1Day
compare_sst <- merge(compare_sst,sst_IUI_1day[,c(1:2)],by = "Date")
#merge dataset daily
compare_sst <- merge(compare_sst,sst_Eilat_IUI_daily[,c(4:5)],by = "Date")

compare_sst <- compare_sst %>% rename("temp_OISST" = "temp",
                                      "mean_1Day"="mean",
                                      "temp_daily"="Temp_avg")
#correlation OISST and mean_1Day
ggplot(compare_sst,aes(x=temp_OISST,y=mean_1Day))+
  geom_point()+
  geom_smooth(method = "lm")+
  stat_regline_equation(label.x=21, label.y=27.5) +
  stat_cor(aes(label=..rr.label..), label.x=21, label.y=27)+
  xlab("OISST")+
  ylab("IUI mean 1 Day")+
  theme_bw()

#correlation OISST and daily
ggplot(compare_sst,aes(x=mean_1Day,y=temp_daily))+
  geom_point()+
  geom_smooth(method = "lm")+
  stat_regline_equation(label.x=21, label.y=27.5) +
  stat_cor(aes(label=..rr.label..), label.x=21, label.y=27)+
  xlab("IUI mean 1Day")+
  ylab("IUI web daily")+
  theme_bw()

#correlation IUI daily and mean 1Day
ggplot(compare_sst,aes(x=temp_OISST,y=temp_daily))+
  geom_point()+
  geom_smooth(method = "lm")+
  stat_regline_equation(label.x=21, label.y=27.5) +
  stat_cor(aes(label=..rr.label..), label.x=21, label.y=27)+
  xlab("OISST")+
  ylab("IUI web daily")+
  theme_bw()

#Timeseries OISST and mean 1Day
ggplot(compare_sst)+
  geom_point(aes(x=Date,y=temp_OISST,color = "OISST"))+
  geom_point(aes(x=Date,y=mean_1Day,color = "IUI mean 1 Day"))+
  scale_x_date(breaks = "6 month",date_labels = "%m-%y")+
  theme_bw()

#Timeseries OISST and IUI daily 
ggplot(compare_sst)+
  geom_point(aes(x=Date,y=temp_OISST,color = "OISST"))+
  geom_point(aes(x=Date,y=temp_daily,color = "IUI daily"))+
  scale_x_date(breaks = "6 month",date_labels = "%m-%y")+
  theme_bw()

#Timeseries mean 1Day and IUI daily 
ggplot(compare_sst)+
  geom_point(aes(x=Date,y=mean_1Day,color = "IUI mean 1 Day"))+
  geom_point(aes(x=Date,y=temp_daily,color = "IUI daily"))+
  scale_x_date(breaks = "6 month",date_labels = "%m-%y")+
  theme_bw()

```

# Detrending
## Jacox 2020
```{r}
# OISST start
red_sst_JACOX <- OISST_data_red %>% filter(lat == OISST_data_red$lat[1] &lon == OISST_data_red$lon[1])
red_sst_JACOX <- red_sst_JACOX %>%  select(-lat,-lon)
red_sst_JACOX <- red_sst_JACOX %>% arrange(t)

# Amazia start
red_sst_JACOX <- IUI_amatzia_daily_temp
red_sst_JACOX <- red_sst_JACOX %>%
  mutate(t = as.Date(Doy - 1, origin = paste0(Year, "-01-01"))) %>%
  select(t, temp = SST)  # Rename columns

# Climatology calculation
clim_period <- c(red_sst_JACOX$t[1],as.Date("2015-12-31"))#edit date to 30 years from beginning if data is available
MHW_climatelogy_tresh <- ts2clm(red_sst_JACOX,climatologyPeriod = clim_period,pctile = 90)
MHW_climatelogy_tresh$anomaly <- MHW_climatelogy_tresh$temp-MHW_climatelogy_tresh$seas
MHW_climatelogy_tresh$anomaly_thresh <- MHW_climatelogy_tresh$thresh-MHW_climatelogy_tresh$seas

# Detrend anomaly
MHW_climatelogy_tresh$num_date <- as.numeric(MHW_climatelogy_tresh$t)
MHW_climatelogy_tresh <- na.omit(MHW_climatelogy_tresh)
warming_lm <- lm(anomaly~num_date,MHW_climatelogy_tresh)
summary(warming_lm)
sjPlot::plot_model(warming_lm,type = 'eff',show.data = T)
sjPlot::plot_model(warming_lm,type = 'diag',show.data = T)
MHW_climatelogy_tresh <- cbind(MHW_climatelogy_tresh,warming_lm$residuals)
names(MHW_climatelogy_tresh)[9] <- "anomaly_detrend"
MHW_climatelogy_tresh$MHW_d <- ifelse(MHW_climatelogy_tresh$anomaly_detrend>MHW_climatelogy_tresh$anomaly_thresh,T,F)
colnames(MHW_climatelogy_tresh)[2] <- "date"
MHW_climatelogy_tresh$date <- as.Date(MHW_climatelogy_tresh$date)
mhws <- Detrend_MHW_days_to_MHWs(MHW_climatelogy_tresh,2,10,9,3,7)

excel <- createWorkbook()
addWorksheet(excel,sheetName = "MHW_Events")
addWorksheet(excel,sheetName = "Climatelogy")
writeData(excel,"MHW_Events",mhws)
writeData(excel,"Climatelogy",MHW_climatelogy_tresh)
saveWorkbook(excel,"Red Sea/MHWs_def/MHW_Amazia_Events_RedSea_Jacox.xlsx")#add description to name according to database used


## Jacox moving baseline explanation
# Extract coefficients
coef_warming <- coef(warming_lm)
slope <- slope <- formatC(coef_warming[2], format = "e", digits = 2)  # Scientific notation
intercept <- round(coef_warming[1], 3)  # Round intercept for readability

# Create the equation text
equation_text <- paste0("y = ", intercept, " + ", slope, "x")


# Aggregate data to monthly means
MHW_monthly <- MHW_climatelogy_tresh %>%
  group_by(year = format(date, "%Y"), month = format(date, "%m")) %>%
  summarize(date = as.Date(paste(year, month, "15", sep = "-")),  # Mid-month,
            seas=mean(seas),
            temp=mean(temp),
            anomaly_threshold=mean(anomaly_thresh),
            anomaly = mean(anomaly, na.rm = TRUE),
            residuals = mean(anomaly_detrend, na.rm = TRUE)) %>%
  ungroup()

# Plot with smoother monthly trends
ggplot(MHW_monthly, aes(y=temp,x = date)) +
  geom_line(aes(color = "Temp"), size = 1.8, alpha = 0.8) +
  # geom_smooth(method = "lm", color = "red", se = TRUE) +  # Linear model with confidence interval
  geom_line(aes(y = seas,color = "Climatology"), size = 1.8, alpha = 0.8) +
  scale_color_manual(values = c("Temp"="black","Climatology"="blue"))+
  theme_minimal()+
  labs(x = "Date",
       y = "Temp",
       color = "Series") +
  theme_minimal()

# Plot Anomalies
ggplot(MHW_monthly, aes(x = date, y = anomaly)) +
  geom_line(aes(color = "Anomaly"), size = 0.3) +  # Plot the 'anomaly' data in black
  # geom_line(aes(y = residuals, color = "Anomaly_detrended"), size = 0.3) +  # Plot 'residuals' data in blue
  geom_line(aes(y = anomaly_threshold, color = "MHW_threshold"),linetype = "dotted",size =1.2) +  # Plot 'anomaly_threshold' data in red
  scale_x_date(limits = as.Date(c("2010-01-01", NA))) +  # Limit x-axis to dates starting from 2010
  scale_color_manual(values = c("MHW_threshold" = "red", "Anomaly" = "black", "Anomaly_detrended" = "blue")) +  # Custom colors for the lines
  labs(x = "Date", y = "SST Anomaly") +
  theme_minimal()


```






## Moving Median detending
```{r}
############### Red Sea #################
red_ts <- ts(OISST_data_red$temp,frequency = 365,start = c(1981,365))
red_mm.ts <- rollmedian(red_ts,93,fill=NA)
red_mm_df <- data.frame(date=OISST_data_red$t,
                    Origin=red_ts,
                    mm=red_mm.ts)
#remove start and end dates without values due to moving median
red_mm_df <- na.omit(red_mm_df)
red_mm_df$subtract <- red_mm_df$Origin - red_mm_df$mm
red_mm_df$doy <- yday(red_mm_df$date)
red_doy_thresh <- red_mm_df %>% 
  select(doy,subtract) %>%
  group_by(doy) %>% 
  summarise("doy_t" = quantile(subtract,0.9,na.rm = T))
red_mm_df <- merge(red_mm_df,red_doy_thresh,by = "doy")
red_mm_df <- red_mm_df %>% arrange(date)
red_mm_df$MHW_d <- ifelse(red_mm_df$subtract>red_mm_df$doy_t,T,F)

#see difference between temp and moving median value
ggplot(red_mm_df,aes(x=date))+
  geom_point(aes(y=Origin),color ="blue")+
  geom_point(aes(y=mm),color ="red")+
  labs(x="Date",
       y="Temp")

#see difference between Moving median anomaly and threshold
ggplot()+
  geom_point(aes(x=red_mm_df$date,red_mm_df$subtract),color = "blue",size =0.4)+
  geom_point(aes(x=red_mm_df$date,red_mm_df$doy_t),size =0.5)+
  xlab("Date")+
  ylab("Anomaly")

#see thresholds in celsius of each day doy
ggplot(red_doy_thresh)+
  geom_point(aes(x=doy,y = doy_t))+
  labs(x="Day of year",
       y="MHW treshold celsius(quantile 90%)")

# mediterranean
# ggplot(med_doy_thresh)+
#   geom_point(aes(x=doy,y = doy_t))+
#   labs(x="Day of year",
#        y="MHW treshold (quantile 90%)")
mhws_31 <- Detrend_MHW_days_to_MHWs(red_mm_df,date_col_num = 2,MHW_day_col_num = 7,residuals_col_num = 5,origin_col_num = 3,threshold_col_num = 6)
hist(mhws$intensity_mean)


######## Mediterranean ############
med_sst_long_1 <- med_sst_long %>% filter(lat == med_sst_long$lat[1] & lon == med_sst_long$lon[1])
med_ts <- ts(med_sst_long_1$sst,frequency = 365,start = c(1981,237)) 
med_mm.ts <- rollmedian(med_ts,93,fill=NA)
med_mm_df <- data.frame(date=med_sst_long_1$date,
                    Origin=med_ts,
                    mm=med_mm.ts)
med_mm_df <- na.omit(med_mm_df)
med_mm_df$subtract <- med_mm_df$Origin - med_mm_df$mm
med_mm_df$doy <- yday(med_mm_df$date)
med_doy_thresh <- med_mm_df %>% 
  select(doy,subtract) %>%
  group_by(doy) %>% 
  summarise("doy_t" = quantile(subtract,0.9,na.rm = T))
med_mm_df <- merge(med_mm_df,med_doy_thresh,by = "doy")
med_mm_df <- med_mm_df %>% arrange(date)
med_mm_df$MHW_d <- ifelse(med_mm_df$subtract>med_mm_df$doy_t,T,F)


excel <- createWorkbook()
addWorksheet(excel,sheetName = "MHW_Events")
addWorksheet(excel,sheetName = "Climatelogy")
writeData(excel,"MHW_Events",mhws_31)
writeData(excel,"Climatelogy",red_mm_df)
saveWorkbook(excel,"Red Sea/Eilat 2017 MHW/MHWs/MHW__detrend_93mm_OISST_Events_RedSea1.xlsx")
```



## Analyze Eilat Linear De-trended Coper & IUI
```{r}
##Coper
red_sst_long <- red_sst_long %>% filter(lat == red_sst_long$lat[1] &lon == red_sst_long$lon[1])
red_sst_long$num_date <- as.numeric(red_sst_long$date)
sst_date_Cop <- ggplot(red_sst_long,aes(x=date,y=sst))+
  geom_point()+
  scale_x_datetime(breaks = "5 year")
warming_lm <- lm(sst~num_date,red_sst_long)
summary(warming_lm)
sjPlot::plot_model(warming_lm,type = 'eff',show.data = T)
sjPlot::plot_model(warming_lm,type = 'diag',show.data = T)

warming_lm_gg <- ggeffects::ggpredict(warming_lm,terms = "num_date[all]")
warming_lm_gg <- cbind(warming_lm_gg,red_sst_long$date)
warming_lm_gg <- cbind(warming_lm_gg,warming_lm$residuals)
colnames(warming_lm_gg)[7:8] <- c("date","residuals")
reisduals_date_Cop <- ggplot(warming_lm_gg,aes(x=date,y=residuals))+
  geom_point()+
  scale_x_datetime(breaks = "5 year")
ggsave(reisduals_date_Cop,"reisduals_date_Cop.png")
hist(warming_lm_gg$residuals)
MHW_thrs <- quantile(warming_lm_gg$residuals,0.9)
warming_lm_gg$MHW_day <- ifelse(warming_lm_gg$residuals>MHW_thrs,T,F)
print(paste("Number of MHW days:", sum(warming_lm_gg$MHW_day==T)))
de_MHW_cop <- Detrend_MHW_days_to_MHWs(warming_lm_gg,7,9,8)
de_MHW_cop$date_start <- as.POSIXct(de_MHW_cop$date_start,origin="1970-01-01")
de_MHW_cop$date_end <- as.POSIXct(de_MHW_cop$date_end,origin="1970-01-01")
de_MHW_cop$date_max <- as.POSIXct(de_MHW_cop$date_max,origin="1970-01-01")


##IUI
sst_IUI_web <- na.omit(sst_IUI_web)
sst_IUI_web$Date <- dmy(sst_IUI_web$Date)
sst_IUI_web$num_date <- as.numeric(sst_IUI_web$Date)
sst_date_IUI <- ggplot(sst_IUI_web,aes(x=Date,y=Temp))+
  geom_point()+
  scale_x_date(breaks = "3 years")
warming_lm_IUI <- lm(Temp~num_date,sst_IUI_web)
summary(warming_lm_IUI)
sjPlot::plot_model(warming_lm_IUI,type = 'eff',show.data = T)
sjPlot::plot_model(warming_lm_IUI,type = 'diag',show.data = T)

warming_lm_IUI_gg <- ggeffects::ggpredict(warming_lm_IUI,terms = "num_date[all]")
warming_lm_IUI_gg <- cbind(warming_lm_IUI_gg,sst_IUI_web$Date)
warming_lm_IUI_gg <- cbind(warming_lm_IUI_gg,warming_lm_IUI$residuals)
colnames(warming_lm_IUI_gg)[7:8] <- c("date","residuals")
warming_lm_IUI_gg$date <- as.Date(warming_lm_IUI_gg$date)
reisduals_date_IUI <- ggplot(warming_lm_IUI_gg,aes(x=date,y=residuals))+
  geom_point()+
  scale_x_date(breaks = "3 years")
hist(warming_lm_IUI_gg$residuals)
MHW_thrs_IUI <- quantile(warming_lm_IUI_gg$residuals,0.9)
warming_lm_IUI_gg$MHW_day <- ifelse(warming_lm_IUI_gg$residuals>MHW_thrs_IUI,T,F)
print(paste("Number of MHW days:", sum(warming_lm_IUI_gg$MHW_day==T)))
de_MHW_IUI <- Detrend_MHW_days_to_MHWs(warming_lm_IUI_gg,7,9,8)
de_MHW_IUI$date_start <- as.Date(de_MHW_IUI$date_start,origin="1970-01-01")
de_MHW_IUI$date_end <- as.Date(de_MHW_IUI$date_end,origin="1970-01-01")
de_MHW_IUI$date_max <- as.Date(de_MHW_IUI$date_max,origin="1970-01-01")


```

# Onset Rates
```{r}
onset_dataset <- OISST_data_red
names(onset_dataset) <- c("date","temp")
onset_dataset$year <- year(onset_dataset$date)
onset_dataset$onset_rate <- NA
for (day_i in 2:length(onset_dataset$date)) {
    onset_dataset$onset_rate[day_i] <- onset_dataset$temp[day_i] - onset_dataset$temp[day_i-1] 
}
hist(onset_dataset$onset_rate)
quantile(onset_dataset$onset_rate,0.99,na.rm = T)

```

# Graphs
```{r}
## Filtering the MHWs to short in the right period
custom_colors <- c("#660000", "#800000", "#990000", "#b30000", "#cc3300", "#e64d00", "#ff8000", "#ffd700")
custom_colors <- rev(custom_colors)
MHW_events_plot_data <- read.xlsx("Red Sea/MHWs_def/MHW_OISST_Events_RedSea_Fix.xlsx",sheet = "Short_MHW_Events_Fixed") #MHW_OISST_Events_RedSea_Fix
MHW_events_plot_data <- MHW_events_plot_data %>%
  select(-reliabaility.level) %>%
  filter(number.of.fish>0)
MHW_events_plot_data$date_start <- as.Date(MHW_events_plot_data$date_start,origin = "1899-12-30")
ggplot(MHW_events_plot_data, aes(x = date_start, y = intensity_max)) +
  geom_lolli(size=2,aes(color = duration)) +
  scale_color_gradientn(colors = custom_colors,name = "Duration \n[Days]")+
  labs(y = expression(paste("Max. intensity [", degree, "C]")), x = NULL) + 
  scale_x_date(breaks = "4 months",date_labels = "%m-%Y")+
  theme_bw()+
  theme(
    axis.title.x = element_text(size = 16),  # Change size of x-axis title
    axis.title.y = element_text(size = 14),  # Change size of y-axis title
    axis.text.x = element_text(size = 14),    # Change size of x-axis text
    axis.text.y = element_text(size = 14),    # Change size of y-axis text
    legend.title = element_text(size = 12),    # Change size of legend title
    legend.text = element_text(size = 12)      # Change size of legend text
  )

ggplot(MHW_events_plot_data, aes(x = date_start, y = duration)) +
  # geom_lolli(size=2,aes(color = intensity_max)) +
  scale_color_gradientn(colors = custom_colors,name = "Max. intensity [C]")+
  labs(x = "Date", 
       y = expression(paste("Duration [Days]")), x = NULL) + 
  scale_x_date(breaks = "4 months",date_labels = "%m-%Y")+
  theme_bw()
  


```
## Fixed baseline explanation
```{r}
## Fixed baseline graphs
# Filter data: keep temp for all years but seas and thresh only after 2011
plot_data <- MHW_climatelogy_tresh %>%
  mutate(seas = ifelse(t >= as.Date("2012-01-01"), seas, NA),
         thresh = ifelse(t >= as.Date("2012-01-01"), thresh, NA))
# Plot
ggplot(plot_data, aes(x = t)) +
  # # Grey rectangle for baseline period (1982-2011)
  # annotate("rect", xmin = as.Date("1982-01-01"), xmax = as.Date("2011-12-31"),
  #          ymin = -Inf, ymax = Inf, fill = "gray23", alpha = 0.2) +
  # 
  # # Baseline period label
  # annotate("text", x = as.Date("1995-01-01"), y = max(plot_data$temp, na.rm = TRUE) + 0.5,
  #          label = "Baseline Period", color = "black", size = 5) +

  # Temperature line (black) - all years
  geom_line(aes(y = temp, color = "Temperature"), size = 1) +

  # Threshold (red) - only after 2011
  geom_line(aes(y = thresh, color = "Threshold"), size = 1) +

  # Climatology (blue) - only after 2011
  geom_line(aes(y = seas, color = "Climatology"), size = 1) +

  # X-axis: Breaks every 5 years
  scale_x_date( breaks = seq(as.Date("2016-01-01"), as.Date("2021-12-31"), by = "1 years"), date_labels = "%Y") +

  # Labels and theme
  labs(x = "Time", y = "Temperature (°C)", color = "Legend") +
  
  # Customize the legend colors
  scale_color_manual(
    values = c("Temperature" = "black", "Threshold" = "red", "Climatology" = "blue")
  ) +
  theme_minimal() +
   # Set the x-axis limit starting from 1/1/2016
  xlim(as.Date("2016-01-01"), as.Date("2021-12-31")) +
  theme(legend.position = "top",
        legend.title = element_blank())

```
